---
title: "Analysis for Growth and calcification responce of the common collector urchin, _Tripneustes gratilla_, to projected climate change: effects of warming and acidification."
author: "Emily Sesno"
output: html_document
---

```{r setup, include=FALSE}

rm(list=ls(all=TRUE))
setwd("/Users/emilysesno/Desktop/Emily/Graduate Work/Urchin_ClimateChange/Data")
results <- read.csv("Urchin_ClimateChange_Data.csv", header=T, na="NA")

library("lsmeans") #post-hoc tests
library("effects") #plot effects of modeling
library("lmtest") #linear mixed modeling
library("lme4") #linear mixed modeling
library("lmerTest") #calculate p-values in models
library("emmeans") #post-hoc tests
library("cowplot") #arrange plots
library("MuMIn")
library("car") #levenes tests, Anova, qqPlots, 
library("tidyverse")
library("stats")
library("plyr")  #splitting, applying, and combining data
library("dplyr")
library("onewaytests") #allows for Welch test with unequal variances (spine lengths)
library("stats")
library("effsize")
library("FSA")
library("rcompanion")
library("broom") #turns model output into data frames


```

## Steps for completing mixed model analysis:
1. Build and run model
2. Check for normality of residuals
3. Check for homogeniety of variance of residuals
4. Look at model summary
5. Run anova to check for significance
6. Run post-hoc

## Growth
Test diameters were not significantly different among all urchins upon initial collection (mean = 7.54 +/- 0.29 mm; n=24; F3,20=0.851; p=0.4825) or after subsequent ramp-up of conditions (mean = 16.12 +/- 0.67 mm; n=24; F3,20=1.0907; p=0.3759). Increased temperatures significantly influenced growth (p = 0.042) while pH did not (p=0.611) (Fig. 3.2). Urchins grown in increased temperatures were 12 % larger than those in ambient conditions, regardless of pH. There was no interaction between temperature and pH (p=0.482) (Fig. 3.2). 

```{r Growth}

### GROWTH MODEL 1: using day 1 as the first official day of experiment when desired environmental conditions were met (24 days after getting urchins. In that 24 days, they were acclimated and then conditions ramped up)


Initial <-
  #seperate out the initial sizes of urchins on day 1, when desired conditions were met.
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column
    filter(Day == "1") %>% 
      #sizes on day 1  
    select("Temperature", "pH", "Diameter", "TankID")


Final <- 
  #seperate out the final sizes of urchins on day 126
   results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column
    filter(Day == "126") %>% 
      #filter for sizes on day 126
    select("Temperature", "pH", "Diameter","TankID")


Growth <- 
  #create column that is growth
  bind_cols(Initial, Final) %>% 
    #binds initial (Diameter) and final (Diameter1) sizes to calculate growth
  mutate(Growth = ((Diameter1 - Diameter)/Diameter*100)) %>% 
    #add column for growth calculation
  select("Growth")


resultsGrow <-
  #table of initial and final size data
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "TankID", "Diam1", "Diam2") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
    filter(Day == "126")
resultsGrow <- bind_cols(resultsGrow, Growth)
  #combines table of initial and final with growth information


#LMM for growth:
modGrow <-
  resultsGrow %>% 
  lmer(Growth~ Temperature * pH + (1|TankID), data=.)

summary(modGrow)
Anova(modGrow, type=2) #Anova{car}. Produces Analysis of Deviance Table (Type II Wald chisquare tests). Calculates type-II or type-III analysis-of-variance tables for model objects produced by lmer in the lme4 package. As implemented here, type-II Wald tests are a generalization of the linear hypotheses used to generate these tests in linear models.
anova(modGrow) #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.

## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modGrow)) #Normal enough?
shapiro.test(residuals(modGrow)) #passes
hist(residuals(modGrow)) #looks normal


# 2. Equal variances
leveneTest(residuals(modGrow)~resultsGrow$Treatment) #Passes
plot(fitted(modGrow),resid(modGrow,type="pearson"),col="blue")

#Interaction plot shows slight interaction
interaction.plot(resultsGrow$Temperature,resultsGrow$pH,resultsGrow$Growth,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(Growth)),)


```


```{r Growth 2 (with day -13 as day 1)}

### GROWTH MODEL 2: If run the same model using day -13 as the first technical day of the experiment. This is the day conditions began to ramp up, so urchins were experiencing gradual increase of stressors.


Initial1 <-
    #seperate out the initial sizes of urchins on day -14, when conditions began to ramp up.
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>%
    filter(TankID != "8t" ) %>% 
        #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>%
      #gather Diam1 and Diam2 into single column
    filter(Day == "-13") %>%  
      #filter for day at -13 to use this as the intial size.
    select("Temperature", "pH", "Diameter", "TankID")


Final1 <- 
      #seperate out the final sizes of urchins on day 126
   results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>%
      #gather Diam1 and Diam2 into single column
    filter(Day == "126") %>% 
      #filter for size on last day of experiment
    select("Temperature", "pH", "Diameter","TankID")


Growth1 <- 
  #create column that is growth
  bind_cols(Initial1, Final1) %>% 
    #binds initial (Diameter) and final (Diameter1) sizes to calculate growth
  mutate(Growth = ((Diameter1 - Diameter)/Diameter*100)) %>%
    #add column for growth calculation
  select("Growth")


resultsGrow1 <-
  #table of initial and final size data
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "TankID", "Diam1", "Diam2") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
    filter(Day == "126")
resultsGrow1 <- bind_cols(resultsGrow1, Growth1)
  #combines table of initial and final with growth information


#LMM for growth:
modGrow1 <- 
  resultsGrow1 %>% 
  lmer(Growth~ Temperature * pH + (1|TankID), data=.)


summary(modGrow1)
Anova(modGrow1, type=2) #Anova{car}. Produces Analysis of Deviance Table (Type II Wald chisquare tests). Calculates type-II or type-III analysis-of-variance tables for model objects produced by lmer in the lme4 package. As implemented here, type-II Wald tests are a generalization of the linear hypotheses used to generate these tests in linear models.
anova(modGrow1) #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.


## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modGrow1)) #More normal than Growth model 1
shapiro.test(residuals(modGrow1)) #passes 
hist(residuals(modGrow1)) #looks normal

# 2. Equal variances
leveneTest(residuals(modGrow1)~resultsGrow1$Treatment) #doesn't pass...
plot(fitted(modGrow1),resid(modGrow1,type="pearson"),col="blue")

#Interaction plot shows interaction
interaction.plot(resultsGrow1$Temperature,resultsGrow1$pH,resultsGrow1$Growth,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(Growth)),)


```


# Calcification
```{r Calcification}

### Model 1: calcification at the tips of spines


resultsTip <- 
  #create data using only the SEM images of the tips of spines.
  results %>% 
  select ("Treatment", "Temperature", "pH", "PartOfSpine", "Chosen", "RatioSEM", "TankID") %>% 
    #selects for desired variables. PartOfSpine is tip or base, Chosen indicates which side of the image (arbitrary) is selected for (aka has no dust)
  drop_na(RatioSEM) %>% 
  filter(Chosen == "yes", PartOfSpine == "Tip")
    #filters only for tips and chosen side. No dusty images to be analyzed.
 

#LMM for calcification at the tips of spines 
modTip <-
  resultsTip %>% 
  lmer(RatioSEM ~ Temperature * pH + (1|TankID), data = .)

summary(modTip)
Anova(modTip, type=2) #Anova{car}. Produces Analysis of Deviance Table (Type II Wald chisquare tests). Calculates type-II or type-III analysis-of-variance tables for model objects produced by lmer in the lme4 package. As implemented here, type-II Wald tests are a generalization of the linear hypotheses used to generate these tests in linear models.
anova(modTip) #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.


## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modTip)) #Normal
shapiro.test(residuals(modTip)) #Pass
hist(residuals(modTip))

# 2. Equal variances
leveneTest(residuals(modTip)~resultsTip$Treatment) #Pass
plot(fitted(modTip),resid(modTip,type="pearson"),col="blue") 

## POST HOC ANALYSIS 
emmTip <- emmeans(modTip, ~Temperature*pH, adjust = "tukey") 
  #Estimated marginal means (Least-squares means)
multcomp::cld(emmTip)
  #create a compact letter display of all pair-wise comparisons
pwpp(emmTip)
  #Constructs a plot of P values associated with pairwise comparisons of estimated marginal means.

#Interaction Tip
interaction.plot(resultsTip$Temperature,resultsTip$pH,resultsTip$RatioSEM,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(Growth)),)

########## Model 2: BASE ###########

resultsBase <-
  results %>% 
  select ("Treatment", "Temperature", "pH", "PartOfSpine", "Chosen", "RatioSEM", "TankID") %>% 
  drop_na(RatioSEM) %>% 
  filter(Chosen == "yes", PartOfSpine == "Base")

modBase <- 
  resultsBase %>% 
  lmer(RatioSEM~ Temperature * pH + (1|TankID), data=.)

summary(modBase) 
Anova(modBase, type=2)

## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modBase)) #Normal
hist(residuals(modBase))
shapiro.test(residuals(modBase)) # Pass
# 2. Equal variances
leveneTest(residuals(modBase)~resultsBase$Treatment) #Pass
plot(fitted(modBase),resid(modBase,type="pearson"),col="blue")

## POST HOC ANALYSIS 
emmBase <- emmeans(modBase, ~Temperature*pH, adjust = "tukey")
multcomp::cld(emmBase)
pwpp(emmBase)

#Interaction Base
interaction.plot(resultsBase$Temperature,resultsBase$pH,resultsBase$RatioSEM,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(RatioSEM)),)

```

# Relative Spine Lengths
```{r, Spine Lengths}

resultsLength <-
  results %>% 
  select ("Treatment", "Temperature", "pH", "RatioLength", "Chosen", "TankID") %>% 
  drop_na(RatioLength) %>% 
  filter(Chosen == "yes")

######  MODEL  ######
modLength <- 
  resultsLength %>% 
  lmer(RatioLength~Temperature*pH + (1|TankID), data=.)

summary(modLength)
Anova(modLength, type=2)

###### ASSSUMPTIONS ######
# 1. Normality of residuals
qqPlot(residuals(modLength)) #Normal
hist(residuals(modLength))
shapiro.test(residuals(modLength)) #Pass
# 2. Equal variances
leveneTest(residuals(modLength)~resultsLength$Treatment) #Pass
plot(fitted(modLength),resid(modLength,type="pearson"),col="blue")

#####  POST HOC ANALYSIS #####
emmLength <- emmeans(modLength, ~Temperature*pH, adjust = "tukey")
multcomp::cld(emmLength)
pwpp(emmLength)

#Interaction Base
interaction.plot(resultsLength$Temperature,resultsLength$pH,resultsLength$RatioLength,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(RatioLength)),)

```

# Dropped Spines
```{r Dropped Spines}

resultsDropped <-
  results %>% 
  select ("Treatment", "Temperature", "pH", "SpineCount", "TankID") %>% 
  drop_na(SpineCount)

## Non parametric ###
kruskal.test(resultsDropped$SpineCount~resultsDropped$Treatment)# SIG*** (p=0.0005563)

### POST HOC Pairwise ######
dunn <- 
  resultsDropped %>% 
  dunnTest(SpineCount ~ Treatment,
           method = "holm", kw = TRUE, data = .)
dunnph <- dunn$res
cldList(P.adj ~ Comparison, data = dunnph, threshold = 0.05)

#Interaction Base
interaction.plot(resultsDropped$Temperature,resultsDropped$pH,resultsDropped$SpineCount,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(SpineCount)),)

```

