---
title: "Ocean acidification reduces skeletal density in spines of the Hawaiian collector urchin, _Tripneustes gratilla_."
author: "Emily E. Sesno, Keisha Bahr, Ariana S. Huffmyer, Judith D. Lemus"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
---

#  Setup  

Set up workspace, load data, and load required packages. 

```{r setup, message=FALSE, warning=FALSE}
rm(list=ls(all=TRUE))

results <- read.csv("Data/Urchin_ClimateChange_Data.csv", header=T, na.strings="NA")

library("lme4") #linear mixed modeling
library("lmtest") #linear mixed modeling
library("lmerTest") #calculate p-values in models
library("car") #levenes tests
library("emmeans") #post-hoc tests
library("lsmeans") #post-hoc tests
library("plotrix") #calculate std.error
library("plyr")  #splitting, applying, and combining data
library("dplyr") 
library("cowplot") #grid plotting and arrange plots
library("effects") #plot effects of modeling
library("MuMIn")
library("car") #levenes tests, Anova, qqPlots, 
library("tidyverse")
library("stats")
library("onewaytests") #allows for Welch test with unequal variances (spine lengths)
library("stats")
library("effsize")
library("FSA")
library("rcompanion")
library("broom") #turns model output into data frames
library("kableExtra") #makes data tables easier to read
library("tidyverse")
library("pwr") #run power tests
```

# Approach:
    1) Build and run model  
    2) Check for normality of residuals  
    3) Check for homogeniety of variance of residuals  
    4) Look at model summary  
    5) Run anova to check for significance of main effects  


# Analysis:

1. Environmental Conditions and Manipulations:
    a. Calculate means
    b. Check environmental manipulations with ANOVAs for treatments
    c. Figure 1: Temp over time
    d. Figure 2: pH over time
    
2. Urchin Diameters: 
    a. Calculate diameter means for:
        i.  Day -24 
        ii. Day -13
        iii.Day 1
        vi. Day 126
        v.  Table 3: Diameter Means
    b. Check for differences in urchin diameters from different tanks on Day 1

3. Response Analyses:
    a. Growth (%)
        i.   Growth Analysis
        ii.  Figure 3: Growth Interaction Plot
    b. Calcification Ratio
        i.   Calcification Analysis at the Tips
        ii.  Calcification Analysis at the Bases
        iii. Figure 4: Calcification Interaction Plot
    c. Spine Loss (count)
        i.   Spine Loss Analysis
        ii.  Figure 5: Spine Loss Interaction Plot
    


## 1. Environmental Conditions and Manipulations:
###  a. Calculate environmental means
##### Table 1: Environmental Means
```{r message=FALSE, warning=FALSE}

EnviroMean <- 
  results %>% 
    select("Day", "Temperature", "pH", "Temp","pHspec", "pCO2out", "AlkTotal") %>%
    filter(Day >= "1", Day <= "123") %>%
    drop_na(Day) %>% 
    group_by(Temperature, pH) %>% 
    summarise(mean_T = mean(Temp, na.rm = T),
              se_T = sd(Temp, na.rm = T)/sqrt(6),
              min_T = min(Temp, na.rm = T), 
              max_T = max(Temp, na.rm = T),
              mean_pH = mean(pHspec, na.rm = T), 
              se_pH = sd(pHspec, na.rm = T)/sqrt(6),
              mean_pCO2 = mean(pCO2out, na.rm = T), 
              se_pCO2 = sd(pCO2out, na.rm = T)/sqrt(6),
              mean_AlkTotal = mean(AlkTotal, na.rm = T),
              se_AlkTotal = sd(AlkTotal, na.rm = T)/sqrt(6))

EnviroMean %>% 
  kbl(caption = "Table 1: Experimental treatments and environmental conditions (mean +/- standard error) of the 126-day experiment.") %>% 
  kable_classic_2() %>% 
  kable_styling(c("striped", "hover"), full_width = F, html_font= "Times New Roman")

```

### b. Check environmental manipulations with ANOVAs
```{r, results=FALSE, message=FALSE, warning=FALSE}

Environmental <- 
  results %>% 
    select("Day", "Temp", "pH", "Temperature","pHspec","Header", "TankID") %>%
    filter(Day >= "1", Day != "126") #use only days from start of experiment when desired conditions were reached
  

#temp between high and ambient temp
tempmod<-aov(Temp ~ Temperature, data=Environmental)
summary(tempmod) #temperature was different between high and ambient treatment

#pH between high and ambient co2
summary(aov(pHspec ~ pH, data=Environmental)) #pH was different between high and ambient co2 treatments



##HEADERS
#temp between headers in high
aov1<- Environmental %>%
  filter(Temperature=="heated")

modaov1<-aov(Temp ~ as.factor(Header), data=aov1)

summary(modaov1) #not different between headers

#temp between headers in ambient temp
aov2<- Environmental %>%
  filter(Temperature=="ambient")

summary(aov(Temp ~ as.factor(Header), data=aov2)) #not different between headers

#pH between headers in acidified
aov3<- Environmental %>%
  filter(pH=="acidified")

summary(aov(pHspec ~ as.factor(Header), data=aov3)) #not different between headers

#pH between headers in ambient pH
aov4<- Environmental %>%
  filter(pH=="ambient")

summary(aov(pHspec ~ as.factor(Header), data=aov4)) #not different between headers


##TANKS
#temp between tanks in high temp
aov1<- Environmental %>%
  filter(Temperature=="heated")

summary(aov(Temp ~ as.factor(TankID), data=aov1)) #not different between heated tanks

#temp between tanks in ambient temp
aov2<- Environmental %>%
  filter(Temperature=="ambient")

summary(aov(Temp ~ as.factor(TankID), data=aov2)) #not different between ambient tanks

#pH between tanks in acidified
aov3<- Environmental %>%
  filter(pH=="acidified")

summary(aov(pHspec ~ as.factor(TankID), data=aov3)) #not different between tanks in acidified

#pH between headers in ambient pH
aov4<- Environmental %>%
  filter(pH=="ambient")

summary(aov(pHspec ~ as.factor(TankID), data=aov4)) #not different between headers

```
##### Table 2: Results of ANOVA for environmental  
Treatment Comparison  | Pr(>F)  
--------------------- |----------
Heated v. Ambient     | <0.001
Acidified v. Ambient  | <0.001
                      |
Ambient temp headers  |  0.998
Heated headers        |  0.842
Ambient pH headers    |  0.129
Acidified headers     |  0.150
                      |
Ambient temp tanks    |  1.000
Heated tanks          |  0.999
Ambient pH tanks      |  0.675
Acidified tanks       |  0.888

Check assumptions of treatment conditions model
```{r, message=FALSE, warning=FALSE, results=FALSE}

#check assumptions
# 1. Normality of residuals
qqPlot(residuals(tempmod)) #very much not normal
shapiro.test(residuals(tempmod)) # nope
hist(residuals(tempmod)) #nope...
```

```{r, message=FALSE, warning=FALSE, results=FALSE}
#check assumptions
# 2. Equal variances
#leveneTest(residuals()~$)
#plot(fitted(),resid(,type="pearson"),col="blue")
```

Check using nonparametric Kruskal-Wallis test
```{r}
# did i set this up right?
kruskal.test(Environmental$Temp~Environmental$Temperature)# **diff
kruskal.test(Environmental$pHspec~Environmental$pH)# **diff

kruskal.test(Environmental$pHspec~Environmental$Temperature)# Not dif
kruskal.test(Environmental$Temp~Environmental$pH)# not dif

```


### c. Figure 1: Temperature over time
```{r, message=FALSE, warning=FALSE, results=FALSE}

tempsummary<-results %>% 
  select("Day", "Temperature", "Temp") %>%
  drop_na(Temp) %>%
  group_by(Day, Temperature) %>%
  mutate(meanT = mean(Temp)) %>%
  group_by(Temperature) %>% 
  mutate(sd = sd(Temp)) %>%
  mutate(se = sd/sqrt(12)) 

tempplot<-ggplot(data=tempsummary, aes(Day, meanT, color = Temperature)) +
    geom_point(size = 2.5, show.legend=FALSE) +
    geom_errorbar(aes(ymin = meanT-se, ymax = meanT+se)) +
    geom_line(size=1.2) +
    geom_vline(xintercept=0) +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="Temperature (°C)", breaks = seq(20,30, by = 1)) +
    scale_shape_discrete(name = NULL,
                         labels = c("Ambient", "High"))+
    scale_color_manual(name = NULL,
                       values = c("gray40", "firebrick3"),
                       labels = c("Ambient", "High")) +
    ggtitle("")+
    theme_classic() +
    theme(plot.margin=unit(c(0.3,0.6,0.3,0.3), "cm"))+
    theme(legend.margin = margin(0), 
          legend.position = c(.98,.8), 
          legend.justification = c("right", "top"),
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14, face="bold"));tempplot

#ggsave(filename="Figures/20200611/TempFig1.png", plot=tempplot, dpi=300, width=7, height=5, units="in")
```

### d. Figure 2: pH over time
```{r message=FALSE, warning=FALSE}

pHsummary<-results %>% 
  select("Day","pH", "pHspec") %>%
  drop_na(pHspec) %>%
  group_by(Day, pH) %>%
  mutate(meanpH = mean(pHspec)) %>%
  group_by(pH, meanpH) %>% 
  mutate(sd = sd(pHspec)) %>%
  mutate(se = sd/sqrt(12)) 
  
  
pHplot<-ggplot(data=pHsummary, aes(Day, meanpH, color = pH)) +
    geom_point(size = 2.5, show.legend=FALSE) +
    geom_errorbar(aes(ymin = meanpH-sd, ymax = meanpH+sd)) +
    geom_line(size=1.2) +
    geom_vline(xintercept=0) +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="pH (Total Scale)", breaks = seq(7.5,8.1, by = .1)) +
    scale_shape_discrete(name = NULL,
                         labels = c("Low pH", "High pH"))+
    scale_color_manual(name = NULL,
                       values = c("skyblue3", "gray40"),
                       labels = c("Acidified", "Ambient"),
                       guide = guide_legend(reverse = TRUE)) +
    ggtitle("")+
    theme_classic() +
    theme(plot.margin=unit(c(0.3,0.6,0.3,0.3), "cm"))+
    theme(legend.margin = margin(0), 
          legend.position = c(.98,.8), 
          legend.justification = c("right", "top"),
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14, face="bold"));pHplot

#ggsave(filename="Figures/20200611/pHFig1.png", plot=pHplot, dpi=300, width=7, height=5, units="in")

```


```{r, results=FALSE, echo=FALSE, warning=FALSE, message=FALSE}

#Make plots in horizontal line with proper labeling of a and b

#treatmentfig1<-plot_grid(tempplot, pHplot, labels = c("a", "b"), ncol=2, nrow=1, rel_heights= c(1,1), rel_widths = c(1,1), label_size = 20, label_y=1, align="h");treatmentfig1

#ggsave(filename="Figures/20200611/TreatmentFig1.png", plot=treatmentfig1, dpi=300, width=16, height=6, units="in")

```



## 2. Urchin Diameters:
### a. Calculate diameter means
#### i.  Day -24
Assemble data for diameters of initial collection of urchins from hatchery and calculate mean
```{r, message=FALSE, warning=FALSE, results=FALSE}

##Initial Collection = Day -24
InitialDiameter <-
  #seperate out the initial sizes of urchins on day -24 after initial collection.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "-24") %>% 
      #sizes on day -24  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


InitialMean <- mean(InitialDiameter$Diameter)
InitialSE <- sd(InitialDiameter$Diameter)/sqrt(23)

InitialMean
InitialSE
```

#### ii. Day -13
Assemble Data for diameters of acclimated urchins before conditioning and calculate mean
```{r, message=FALSE, warning=FALSE, results=FALSE}
##After acclimating before ramp up = Day-13

AcclimatedDiameter <-
  #seperate out the initial sizes of urchins on day -24, when desired conditions were met.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "-13") %>% 
      #sizes on day -24  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


AcclimatedMean <- mean(AcclimatedDiameter$Diameter)
AcclimatedSE <- sd(AcclimatedDiameter$Diameter)/sqrt(23) 

AcclimatedMean
AcclimatedSE
```


#### iii. Day 1
Assemble Data for diameters on day 1 of experiment when desired conditions were reached
```{r, message=FALSE, warning=FALSE, results=FALSE}
##After ramp up and desired conditions are reached to begin experiment = Day 1

Day1Diameter <-
  #seperate out the initial sizes of urchins on day 1, when desired conditions were met.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "1") %>% 
      #sizes on day 1  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


Day1Mean <- mean(Day1Diameter$Diameter)
Day1SE <- sd(Day1Diameter$Diameter)/sqrt(23)

Day1Mean
Day1SE

```

#### iv. Day 126
Assemble Data for diameters on day 126 of experiment after full growth
```{r, message=FALSE, warning=FALSE, results=FALSE}
##After ramp up and desired conditions are reached to begin experiment = Day 1

Day126Diameter <-
  #seperate out the initial sizes of urchins on day 1, when desired conditions were met.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "126") %>% 
      #sizes on day 1  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


Day126Mean <- mean(Day126Diameter$Diameter)
Day126SE <- sd(Day126Diameter$Diameter)/sqrt(23)

Day126Mean
Day126SE

```

##### v. Table 3: Urchin diameter means 
Day  | Mean Urchin Diameter (mm) | se
-----|---------------------------| -----
-24  | 7.53                      | 0.15
-13  | 10.38                     | 0.23
1    | 16.03                     | 0.36
126  | 70.52                     | 1.43

### b. Results of linear mixed model effect on diameter (day 1)
```{r, message=FALSE, warning=FALSE}
Day1Mod <- lmer(Diameter ~ Temperature*pH + (1|TankID), data=Day1Diameter)
anova(Day1Mod, type = 2)
```

Check assumptions of diameter model (Day 1)
```{r, message=FALSE, warning=FALSE, results=FALSE}
#check assumptions
# 1. Normality of residuals
qqPlot(residuals(Day1Mod)) #normal
shapiro.test(residuals(Day1Mod)) #passes
hist(residuals(Day1Mod)) #normal

```
```{r, message=FALSE, warning=FALSE, results=FALSE}
#check assumptions
# 2. Equal variances
leveneTest(residuals(Day1Mod)~InitialDiameter$Treatment) #Passes
plot(fitted(Day1Mod),resid(Day1Mod,type="pearson"),col="blue")
```


## 3. Response Analysis
### a. Growth
#### i. Growth Analysis
Assemble data and calculate means using day 1 as initial diameter and calculate means
```{r message=FALSE, warning=FALSE}
### GROWTH MODEL 1: using day 1 as the first official day of experiment when desired environmental conditions were met (24 days after getting urchins. In that 24 days, they were acclimated and then conditions ramped up)

Initial <-
  #seperate out the initial sizes of urchins on day 1, when desired conditions were met to standardize growth to this measurement.
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "1") %>% 
      #sizes on day 1  
    select("Temperature", "pH", "Diameter", "TankID")%>%
    rename(InitialDiameter=Diameter)

  
Final <- 
  #seperate out the final sizes of urchins on day 126
   results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column
    filter(Day == "126") %>% 
      #filter for sizes on day 126
    select("Temperature", "pH", "Diameter","TankID")%>%
    rename(FinalDiameter=Diameter)


resultsGrow <- 
  #create column that is growth
  full_join(Initial, Final) %>% 
    #binds initial (Diameter) and final (Diameter1) sizes to calculate growth
  mutate(Growth = ((FinalDiameter - InitialDiameter)/InitialDiameter*100)) %>% 
    #add column for growth calculation
  select("Temperature", "pH", "TankID", "Growth")

##WE DONT NEED THIS CODE ANYMORE, THE FULL JOIN ABOVE DOES ALL THIS IN A MORE SECURE WAY 
#resultsGrow <-
  #table of initial and final size data
  #results %>% 
    #select("Day", "Treatment", "Temperature", "pH", "TankID", "Diam1", "Diam2") %>% 
    #drop_na(Diam1, Diam2) %>% 
    #gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
    #filter(Day == "126")
#resultsGrow <- full_join(resultsGrow, Growth)
#combines table of initial and final with growth information

#Figure out means of each treatment group
resultsGrow %>% 
  dplyr::group_by(Temperature, pH) %>% 
  dplyr::summarise(mean = mean(Growth), s.e. = se(Growth), n=length(Growth)) %>% 

  kbl(caption = "Table 6: Growth means (%) +/- se after 126 days") %>% 
  kable_classic_2() %>% 
  kable_styling(c("striped", "hover"), full_width = F, html_font= "Times New Roman")

```

Growth ANOVA results
```{r, message=FALSE, warning=FALSE}
#LMM for growth:
modGrow <-
  resultsGrow %>% 
  aov(Growth~ Temperature * pH, data=.)

summary(modGrow)

```

Check assumptions for growth model
```{r, message=FALSE, warning=FALSE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modGrow)) 
shapiro.test(residuals(modGrow)) #passes
hist(residuals(modGrow)) #looks normal

```

```{r, message=FALSE, warning=FALSE}
## ASSSUMPTIONS
# 2. Equal variances
leveneTest(residuals(modGrow)~resultsGrow$Temperature) #Passes
leveneTest(residuals(modGrow)~resultsGrow$pH) #doesn't pass
plot(fitted(modGrow),resid(modGrow,type="pearson"),col="blue")
```
Now, calculate effect size for difference in growth by temperature.  

```{r}
library("effsize")
library("rstatix")

#calculate cohens D effect size
cohens_d(Growth ~ Temperature, var.equal = FALSE, ref.group="ambient", data=resultsGrow)

```
From this, we see that we have a large effect size between temperature treatments (-0.86). 

Confirm that we have the sapmle size necessary to detect a large effect. Power analysis for test
```{r, message=FALSE, warning=FALSE}

#calculate test power first given known variables  
pwr.anova.test(k=4, n=6, f=0.8, sig.level = 0.05) #balanced one way anova

#k=Number of groups - 4 treatments
#n=Number of observations (per group) - 6 
#f=Effect size - 0.8 based on above calculation
#sig.level=Significance level (Type I error probability)

#solve for power=Power of test (1 minus Type II error probability)

#this last test identified power as 0.86, now look for the sample size we would need to detect a small (0.2), medium (0.5) or large (0.8) effect size. 

pwr.anova.test(k=4, f=0.2, sig.level = 0.05, power=0.86) #n needed for small effect, would need n=79 urchins
pwr.anova.test(k=4, f=0.5, sig.level = 0.05, power=0.86) #n needed for small effect, would need n=13 urchins 
pwr.anova.test(k=4, f=0.8, sig.level = 0.05, power=0.86) #n needed for small effect, would need n=6 urchins

#pwr.f2.test(u=1, v=42, f=0.7, sig.level = 0.05) #general linear model

#u	= degrees of freedom for numerator
#v	= degrees of freedom for denominator
#f2	= effect size
#sig.level	= Significance level (Type I error probability)
#power	= Power of test (1 minus Type II error probability)

```

This power analysis for growth shows us that in order to detect a small effect, we would need 79 urchins. To detect a medium effect, we would need 13 urchins. To detect a large effect, we would need 6 urchins. Given that we have n=6 urchins per group, we would have the statistical power to detect a large effect that was detected. This validates our tst power. 

We can repeat this approach for each response.  

### ii. Growth Interaction Plot 
```{r, message=FALSE, warning=FALSE}

growthsummary<-resultsGrow %>% 
  #filter(Day==126) %>%
  select("pH", "Temperature","Growth") %>%
  drop_na(Growth) %>% 
  mutate(meanPct = Growth) %>%
  group_by(pH, Temperature) %>%
  summarise(mean = mean(meanPct), 
            N = length(meanPct),
            se = std.error(meanPct))
   
  
growthplot<-ggplot(data=growthsummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Total Growth (%)")+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    geom_text(x=1.5, y=125, label="p(pH)=0.466", size=6, color="darkgray") + 
    geom_text(x=1.5, y=100, label="p(Temperature)=0.006", size=6, color="black") + 
    geom_text(x=1.5, y=75, label="p(Interaction)=0.303", size=6, color="darkgray") + 
    ylim(0,450)+
    theme_classic()+
    theme(legend.position = "right", 
          legend.background = element_blank(),
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));growthplot

ggsave(filename="Figures/20200611/GrowthFig.png", plot=growthplot, dpi=300, width=8, height=6, units="in")
```

### b. Growth over time analysis (AH ADDED DAY TO THIS MODEL)  
```{r, message=FALSE, warning=FALSE}

resultsGrowTime <-  #create table of growth in treatments
  results %>% 
  select("Day", "TankID", "Treatment","Temperature","pH","Growth") %>%
  drop_na(Growth) %>% 
  group_by(Day, Treatment) %>%
  mutate(se = (std.error(Growth))*100) %>%
  mutate(meanPct = (mean(Growth)*100))

modelGrowTime <- lmer(Growth~Temperature*pH*Day + (1|TankID), data=resultsGrowTime)
summary(modelGrowTime) #generate results
anova(modelGrowTime, type=2)

```


#### i. Growth over time plot across all treatments
```{r, message=FALSE, warning=FALSE}

#### Growth Across Treatments
results %>% 
  select("Day", "Treatment","Temperature","pH","Growth") %>%
  drop_na(Growth) %>% 
  group_by(Day, Treatment) %>%
  mutate(se = (std.error(Growth))*100) %>%
  mutate(meanPct = (mean(Growth)*100)) %>% 

  
ggplot(aes(Day, meanPct, shape = Treatment, color = Treatment)) +
    geom_point(size = 2.5) +
    geom_line() +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="Growth (%)", breaks = seq(0,400, by = 25)) +
    geom_errorbar(aes(ymin = meanPct-se, ymax = meanPct+se, width = 1.5)) +
    theme_classic() + 
    scale_shape_discrete(name = NULL,
                         labels = c("Amb T, Amb pH", "Amb T, Low pH", "High T, Amb pH", "High T, Low pH")) +
    scale_colour_manual(name = NULL,
                        values = c("gray40", "skyblue3", "firebrick3","mediumpurple3"),
                        labels = c("Amb T, Amb pH", "Amb T, Low pH", "High T, Amb pH", "High T, Low pH")) +
    theme(legend.margin = margin(0), 
          legend.position = c(.1, .98), 
          legend.justification = c("left", "top"),
          legend.background = element_blank())

ggsave(filename="Figures/20200611/GrowthTreatments.png", dpi=300, width=8, height=6, units="in")

```

#### ii. Growth over time plot by temperature
```{r}
results %>% 
  select("Day","Temperature","Growth") %>%
  drop_na(Growth) %>% 
  group_by(Day, Temperature) %>%
  mutate(se = (std.error(Growth))*100) %>%
  mutate(meanPct = (mean(Growth)*100)) %>% 

  
  ggplot(aes(Day, meanPct, shape = Temperature, color = Temperature)) +
    geom_point(size = 2.5) +
    geom_line() +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="Growth (%)", breaks = seq(0,400, by = 25)) +
    geom_errorbar(aes(ymin = meanPct-se, ymax = meanPct+se, width = 1.5)) +
    theme_classic() +
    scale_colour_manual(name = NULL, 
                        values = c("gray40", "firebrick3"), 
                        labels = c("Ambient Temperture", "High Temperature")) +
    scale_shape_discrete(name = NULL, 
                       labels = c("Ambient Temperture", "High Temperature")) +
    theme(legend.margin = margin(0), 
          legend.position = c(.03, .93), 
          legend.justification = c("left", "top")) + 
    annotate("text", x = 0, y = 375, label = "(a)") +
    annotate("text", x = 130, y = 359, label = "*", size = 7)
```

#### iii. Growth over time plot by pH
```{r}
results %>% 
  select("Day","pH","Growth") %>%
  drop_na(Growth) %>% 
  group_by(Day, pH) %>%
  mutate(se = (std.error(Growth))*100) %>%
  mutate(meanPct = (mean(Growth)*100)) %>% 

  ggplot(aes(Day, meanPct, shape = pH, color = pH)) +
    geom_point(size = 2.5) +
    geom_line() +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="Growth (%)", breaks = seq(0,400, by = 25)) +
    geom_errorbar(aes(ymin = meanPct-se, ymax = meanPct+se, width = 1.5)) +
    theme_classic() +
    scale_colour_manual(name = NULL, 
                        values = c("skyblue3", "gray40"), 
                        labels = c("Low pH", "Ambient pH")) +
    scale_shape_discrete(name = NULL,
                         labels = c("Low pH", "Ambient pH")) +
    theme(legend.margin = margin(0), 
          legend.position = c(.03, .93), 
          legend.justification = c("left", "top")) + 
    annotate("text", x = 0, y = 370, label = "(b)")
    

```


### b. Calcification Ratio
#### i. Calcification Ratio Analysis at the tips

Assemble data for chosen spine tips (distal end)
```{r, message=FALSE, warning=FALSE}
### Model 1: calcification at the tips of spines

resultsTip <- 
  #create data using only the SEM images of the tips of spines.
  results %>% 
  select ("Treatment", "Temperature", "pH", "PartOfSpine", "Chosen", "RatioSEM", "TankID") %>% 
    #selects for desired variables. PartOfSpine is tip or base, Chosen indicates which side of the image (arbitrary) is selected for (aka has no dust)
  drop_na(RatioSEM) %>% 
  filter(Chosen == "yes", PartOfSpine == "Tip")
    #filters only for tips and chosen side. No dusty images to be analyzed.
```

Results for spine tips with linear mixed model
```{r, message=FALSE, warning=FALSE}
#LMM for calcification at the tips of spines 
modTip <-
  resultsTip %>% 
  lmer(RatioSEM ~ Temperature * pH + (1|TankID), data = .)

summary(modTip)

anova(modTip, type="II") #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type II Analysis of Variance Table with Satterthwaite's method.
```

Check assumptions for calcification at tip model.  
```{r, message=FALSE, warning=FALSE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modTip)) #Normal
shapiro.test(residuals(modTip)) #Pass
hist(residuals(modTip))

# 2. Equal variances
leveneTest(residuals(modTip)~resultsTip$Treatment) #Pass
plot(fitted(modTip),resid(modTip,type="pearson"),col="blue") 
```


#### ii. Calcification Ratio Analysis at the base
Assemble data for chosen spine bases (proximal end)
```{r, message=FALSE, warning=FALSE}

### Model 2: calcification in the base of the spines

resultsBase <-
   #create data using only the SEM images of the base of spines.
  results %>% 
  select ("Treatment", "Temperature", "pH", "PartOfSpine", "Chosen", "RatioSEM", "TankID") %>% 
    #selects for desired variables. PartOfSpine is tip or base, Chosen indicates which side of the image (arbitrary) is selected for (aka has no dust)
  drop_na(RatioSEM) %>% 
  filter(Chosen == "yes", PartOfSpine == "Base")
    #filters only for tips and chosen side. No dusty images to be analyzed.
```

Results for spine bases with linear mixed model
```{r, message=FALSE, warning=FALSE}
#LMM for calcification at the tips of spines 
modBase <- 
  resultsBase %>% 
  lmer(RatioSEM~ Temperature * pH + (1|TankID), data=.)

summary(modBase) 
anova(modBase, type="II") #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.
```

Check assumptions for calcification at base model. 
```{r, message=FALSE, warning=FALSE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modBase)) #Normal
shapiro.test(residuals(modBase)) # Pass
hist(residuals(modBase))

# 2. Equal variances
leveneTest(residuals(modBase)~resultsBase$Treatment) #Pass
plot(fitted(modBase),resid(modBase,type="pearson"),col="blue")
```


#### iii. Figure 4: Calcification Ratio Interaction Plot at tips and bases
Make Interaction plots for calcification at the base and the tip
```{r, message=FALSE, warning=FALSE}
#### Tip and base calcification ratios

#base
basesummary<-results %>% 
  select("Temperature","pH","RatioSEM", "PartOfSpine") %>%
  filter(PartOfSpine=="Base")%>%
  drop_na(RatioSEM) %>% 
  group_by(pH, Temperature, PartOfSpine) %>%
  summarize(mean = mean(RatioSEM), se = std.error(RatioSEM), N = length(RatioSEM))
  
baseplot<- ggplot(data=basesummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Proximal Calcification Ratio")+
    ylim(1,3.5)+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    geom_text(x=1.5, y=1.6, label="p(pH)=0.002", size=6, color="black") + 
    geom_text(x=1.5, y=1.4, label="p(Temperature)=0.16", size=6, color="darkgray") + 
    geom_text(x=1.5, y=1.2, label="p(Interaction)=0.54", size=6, color="darkgray") + 
    theme_classic()+
    theme(legend.position = "right", 
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));baseplot

#ggsave("CalcRatiobase.png", plot=baseplot, path = "/Users/emilysesno/Desktop/R_Analysis/R_Analysis/output",  width = 7, height = 5)  

ggsave(filename="Figures/20200611/BaseCalcificationFig.png", plot=baseplot, dpi=300, width=8, height=6, units="in")

#tip
tipsummary<-results %>% 
  select("Temperature","pH","RatioSEM", "PartOfSpine") %>%
  filter(PartOfSpine=="Tip")%>%
  drop_na(RatioSEM) %>% 
  group_by(pH, Temperature, PartOfSpine) %>%
  summarize(mean = mean(RatioSEM), se = std.error(RatioSEM), N = length(RatioSEM))

tipplot<-ggplot(data=tipsummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Distal Calcification Ratio")+
    ylim(1,3.5)+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    geom_text(x=1.5, y=3.0, label="p(pH)=0.44", size=6, color="darkgray") + 
    geom_text(x=1.5, y=2.8, label="p(Temperature)=0.39", size=6, color="darkgray") + 
    geom_text(x=1.5, y=2.6, label="p(Interaction)=0.09", size=6, color="darkgray") + 
    theme_classic()+
    theme(legend.position = "none", #removed the legend for the tip, so when we align them horizontally we only have one legend. But need to adjust the size so it is the same as base
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));tipplot


#ggsave(filename="Figures/20200611/TipCalcificationFig.png", plot=tipplot, dpi=300, width=8, height=6, units="in")


```

Combine both interaction plots at base and tip to be in horizontal line
```{r, message=FALSE, warning=FALSE, results = FALSE, echo=FALSE}

#Make plots of calcification at tip and base in horizontal line with proper labeling of a and b
#removed legend on the tip, so they can share the same one. But now they are different sizes...

#calcificationfig1<-plot_grid(tipplot, baseplot, labels = c("a", "b"), ncol=2, nrow=1, rel_heights= c(1,1), rel_widths = c(0.8,1), label_size = 20, label_y=1, align="h");calcificationfig1

#ggsave(filename="Figures/20200611/CalcificationFig1.png", plot=calcificationfig1, dpi=300, width=16, height=6, units="in")

```


### c. Spine Loss
#### i. Spine Loss Analysis:
Assemble data for dropped spines  
```{r, message=FALSE, warning=FALSE}

## Number of loose spines counted at the bottom of tanks at the end of the experiment. These were either broken or shed. Note: these spines were not collected, just counted, so can not determine whether they were broken or shed.

resultsDropped <-
  #create data using only the needed variables.
  results %>% 
  select ("Treatment", "Temperature", "pH", "SpineCount", "TankID") %>% 
  drop_na(SpineCount) %>% 
  group_by(pH) %>% 
  mutate(mean = mean(SpineCount),
         se = std.error(SpineCount))
```

Analyse dropped spines with linear mixed effect model.  
```{r, message=FALSE, warning=FALSE}
##LM of dropped spines - can't make it LMM because error: number of levels of each grouping factor must be < number of observations - only have one count for each TankID.
modDropped <- 
  resultsDropped %>% 
  lm(SpineCount~ Temperature * pH, data=.)

summary(modDropped)
anova(modDropped) #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.
```

Check assumptions for model for dropped spines
```{r, message=FALSE, warning=FALSE}
##ASSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modDropped)) #not normal
shapiro.test(residuals(modDropped)) #faaaiiilll
hist(residuals(modDropped))

# 2. Equal variances
leveneTest(residuals(modDropped)~resultsDropped$Treatment) # barely pass
plot(fitted(modDropped),resid(modDropped,type="pearson"),col="blue")
```

Analysis for dropped spines does not meet assumptions. Attempt a square root transformation.  
```{r, message=FALSE, warning=FALSE}
##DOES NOT MEET ASSUMPTIONS so ATTEMPT AT Transformation:
resultsDropped$tdata<-(resultsDropped$SpineCount)^1/2
  # transformation
modDropped1 <- lm(tdata~ Temperature * pH, data=resultsDropped)
  # run lm model of transformed data

###ASSUMPTIONS for Transformation
# 1. Normality of residuals
qqPlot(residuals(modDropped1)) #not normal
shapiro.test(residuals(modDropped1)) #fail
hist(residuals(modDropped1))
# 2. equal variances
bartlett.test(residuals(modDropped1)~resultsDropped$Treatment) #fail
plot(fitted(modDropped1),resid(modDropped1,type="pearson"),col="blue")


summary(modDropped1)
Anova(modDropped1, type=2) #test for significance of model
```

Transfomraiton (of any usual kind) of dropped spines was not successful, so we use a non parametric Kruskal Wallis Test.  
```{r, results = FALSE, message=FALSE, warning=FALSE}
### Transformations not work, so use non parametric:
kruskal.test(resultsDropped$SpineCount~resultsDropped$Treatment)# SIG*** (p=0.0005563)

kruskal.test(resultsDropped$SpineCount~resultsDropped$Temperature)# 
kruskal.test(resultsDropped$SpineCount~resultsDropped$pH)#
```

Conduct Dunn Post Hoc Test for dropped spines
```{r, message=FALSE, warning=FALSE}
## POST HOC Pairwise
dunn <- 
  resultsDropped %>% 
  dunnTest(SpineCount ~ Treatment,
           method = "holm", kw = TRUE, data = .)
dunnph <- dunn$res
cldList(P.adj ~ Comparison, data = dunnph, threshold = 0.05)
```

#### ii. Figure 5: Spine Loss Interaction Plot
```{r, message=FALSE, warning=FALSE}
droppedsummary<-results %>% 
  select("Temperature","pH","SpineCount") %>%
  drop_na(SpineCount) %>% 
  group_by(pH, Temperature) %>%
  summarize(mean = mean(SpineCount), se = std.error(SpineCount), N = length(SpineCount))

  dropplot<-ggplot(data=droppedsummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Dropped Spines")+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    ylim(0,30) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    #geom_text(x=1.5, y=11, label="p(pH)<0.001", size=6, color="black") + 
    #geom_text(x=1.5, y=9, label="p(Temperature)=0.91", size=6, color="darkgray") + 
    #geom_text(x=1.5, y=7, label="p(Interaction)=0.83", size=6, color="darkgray") +
    theme_classic()+
    theme(legend.position = "right", 
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));dropplot

#ggsave(filename="Figures/20200611/DroppedSpinesFig.png", plot=dropplot, dpi=300, width=8, height=6, units="in")
```

