---
title: "Analysis for Growth and calcification response of the common collector urchin, _Tripneustes gratilla_, to projected climate change: effects of warming and acidification."
author: "Emily E. Sesno, Keisha Bahr, Ariana S. Huffmyer, Judith D. Lemus"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
---

```{r setup, message=FALSE, warning=FALSE}

rm(list=ls(all=TRUE)) 
results <- read.csv("Data/Urchin_ClimateChange_Data.csv", header=T, na="NA")

library("lsmeans") #post-hoc tests
library("effects") #plot effects of modeling
library("lmtest") #linear mixed modeling
library("lme4") #linear mixed modeling
library("lmerTest") #calculate p-values in models
library("emmeans") #post-hoc tests
library("cowplot") #arrange plots
library("MuMIn")
library("car") #levenes tests, Anova, qqPlots, 
library("tidyverse")
library("stats")
library("plyr")  #splitting, applying, and combining data
library("dplyr")
library("onewaytests") #allows for Welch test with unequal variances (spine lengths)
library("stats")
library("effsize")
library("FSA")
library("rcompanion")
library("broom") #turns model output into data frames
library("plotrix") #Calculates standard error of the mean
```

## Analysis approach:  
1. Build and run model  
2. Check for normality of residuals  
3. Check for homogeniety of variance of residuals  
4. Look at model summary  
5. Run anova to check for significance of main effects  
6. Run post-hoc to test hypotheses  


## Data Assembly and Results:

### Diameter Means

#### Day -24
##### Assemble data for diameters of initial collection of urchins from hatchery (see code)
```{r InitialDiameter=TRUE}
##Initial Collection = Day -24
InitialDiameter <-
  #seperate out the initial sizes of urchins on day -24, when desired conditions were met.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "-24") %>% 
      #sizes on day -24  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


InitialMean <- mean(InitialDiameter$Diameter)
InitialSE <- std.error(InitialDiameter$Diameter)

```

##### Results of linear mixed model effect on diameter (day -24)
```{r InitialResults=TRUE}
InitialMod <- lmer(Diameter ~ Temperature*pH + (1|TankID), data=InitialDiameter)
anova(InitialMod, type = 2)
```

##### Check assumptions of diameter model (day -24)
```{r InitialResultsAssumptions=TRUE}
#check assumptions
# 1. Normality of residuals
qqPlot(residuals(InitialMod)) 
shapiro.test(residuals(InitialMod)) #no pass
hist(residuals(InitialMod)) #eh


# 2. Equal variances
leveneTest(residuals(InitialMod)~InitialDiameter$Treatment) #No pass
plot(fitted(InitialMod),resid(InitialMod,type="pearson"),col="blue")

```


#### Day -13
##### Assemble Data for diameters of acclimated urchins before conditioning (see code)
```{r AcclimatedDiameter=TRUE}
##After acclimating before ramp up = Day-13

AcclimatedDiameter <-
  #seperate out the initial sizes of urchins on day -24, when desired conditions were met.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "-13") %>% 
      #sizes on day -24  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


AcclimatedMean <- mean(AcclimatedDiameter$Diameter)
AcclimatedSE <- std.error(AcclimatedDiameter$Diameter)

```

##### Results of linear mixed model effect on diameter (day -13)
```{r AcclimatedResults=TRUE}
AcclimatedMod <- lmer(Diameter ~ Temperature*pH + (1|TankID), data=AcclimatedDiameter)
anova(AcclimatedMod, type = 2)

```

##### Check assumptions of diameter model (day -13)
```{r AcclimatedResults=TRUE}
#check assumptions
# 1. Normality of residuals
qqPlot(residuals(AcclimatedMod)) #not super normal but... see below
shapiro.test(residuals(AcclimatedMod)) #passes
hist(residuals(AcclimatedMod)) #looks normalish


# 2. Equal variances
leveneTest(residuals(AcclimatedMod)~InitialDiameter$Treatment) #Passes
plot(fitted(AcclimatedMod),resid(AcclimatedMod,type="pearson"),col="blue")

```


#### Day 1
##### Assemble Data for diameters on day 1 of experiment when desired conditions were reached (see code)
```{r Day1Diameter=TRUE}
##After ramp up and desired conditions are reached to begin experiment = Day 1

Day1Diameter <-
  #seperate out the initial sizes of urchins on day 1, when desired conditions were met.
  results %>% 
    select("Day", "Diam1", "Diam2", "TankID", "Temperature", "pH", "Treatment") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "1") %>% 
      #sizes on day 1  
    select("TankID", "Diameter", "Temperature", "pH", "Treatment")


Day1Mean <- mean(Day1Diameter$Diameter)
Day1SE <- std.error(Day1Diameter$Diameter)


```

##### Results of linear mixed model effect on diameter (day 1)
```{r Day1Results=TRUE}
Day1Mod <- lmer(Diameter ~ Temperature*pH + (1|TankID), data=Day1Diameter)
anova(Day1Mod, type = 2)
```

##### Check assumptions of diameter model (Day 1)
```{r Day1Results=TRUE}
#check assumptions
# 1. Normality of residuals
qqPlot(residuals(AcclimatedMod)) #not super normal but... see below
shapiro.test(residuals(AcclimatedMod)) #passes
hist(residuals(AcclimatedMod)) #looks normalish


# 2. Equal variances
leveneTest(residuals(AcclimatedMod)~InitialDiameter$Treatment) #Passes
plot(fitted(AcclimatedMod),resid(AcclimatedMod,type="pearson"),col="blue")

```


### Growth  


#### Initial as Day -13
##### Assemble data and calculate means using day -13 as initial diameter (see code)
```{r, results=FALSE}

### GROWTH MODEL 2: If run the same model using day -13 as the first technical day of the experiment. This is the day conditions began to ramp up, so urchins were experiencing gradual increase of stressors.


Initial1 <-
    #seperate out the initial sizes of urchins on day -14, when conditions began to ramp up.
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>%
    filter(TankID != "8t" ) %>% 
        #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>%
      #gather Diam1 and Diam2 into single column
    filter(Day == "-13") %>%  
      #filter for day at -13 to use this as the intial size.
    select("Temperature", "pH", "Diameter", "TankID")

  mean(Initial1$Diameter)
  std.error(Initial1$Diameter)


Final1 <- 
      #seperate out the final sizes of urchins on day 126
   results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>%
      #gather Diam1 and Diam2 into single column
    filter(Day == "126") %>% 
      #filter for size on last day of experiment
    select("Temperature", "pH", "Diameter","TankID")


Growth1 <- 
  #create column that is growth
  bind_cols(Initial1, Final1) %>% 
    #binds initial (Diameter) and final (Diameter1) sizes to calculate growth
  mutate(Growth = ((Diameter1 - Diameter)/Diameter*100)) %>%
    #add column for growth calculation
  select("Growth")


resultsGrow1 <-
  #table of initial and final size data
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "TankID", "Diam1", "Diam2") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
    filter(Day == "126")
resultsGrow1 <- bind_cols(resultsGrow1, Growth1)
  #combines table of initial and final with growth information


#Figure out means of each treatment group
resultsGrow1 %>% 
  dplyr::group_by(Temperature, pH) %>% 
  dplyr::summarise(mean = mean(Growth), s.e. = se(Growth))

```

##### Results with linear mixed model effect on Growth (Day -13 as initial).  
```{r, results=TRUE}
#LMM for growth:
modGrow1 <- 
  resultsGrow1 %>% 
  lmer(Growth~ Temperature * pH + (1|TankID), data=.)


summary(modGrow1)
anova(modGrow1, type="II") #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.

```

##### Check assumptions of growth model (day -13 as initial)  
```{r, results=TRUE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modGrow1)) 
shapiro.test(residuals(modGrow1)) #passes 
hist(residuals(modGrow1)) #looks normal

# 2. Equal variances
leveneTest(residuals(modGrow1)~resultsGrow1$Treatment) #doesn't pass...
plot(fitted(modGrow1),resid(modGrow1,type="pearson"),col="blue")

```

##### Display exploratory interaction plot for growth model (day -13 as initial).  
```{r, results=TRUE}

#Interaction plot shows interaction
interaction.plot(resultsGrow1$Temperature,resultsGrow1$pH,resultsGrow1$Growth,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(Growth)),)

```


#### Day 1
##### Assemble data and calculate means using day 1 as initial diameter and calculate means (see code) 
```{r Growth=TRUE}
### GROWTH MODEL 1: using day 1 as the first official day of experiment when desired environmental conditions were met (24 days after getting urchins. In that 24 days, they were acclimated and then conditions ramped up)

Initial <-
  #seperate out the initial sizes of urchins on day 1, when desired conditions were met to standardize growth to this measurement.
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>%
    filter(TankID != "8t" ) %>% 
      #remove urchin number 8 which died halfway through
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column in order to make number of levels of each grouping factor < number of observations (as per error message previously recieved)
    filter(Day == "1") %>% 
      #sizes on day 1  
    select("Temperature", "pH", "Diameter", "TankID")

#Mean diameters on day 1
  mean(Initial$Diameter)
  std.error(Initial$Diameter)

  
Final <- 
  #seperate out the final sizes of urchins on day 126
   results %>% 
    select("Day", "Treatment", "Temperature", "pH", "Diam1", "Diam2", "TankID") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
      #gather Diam1 and Diam2 into single column
    filter(Day == "126") %>% 
      #filter for sizes on day 126
    select("Temperature", "pH", "Diameter","TankID")


Growth <- 
  #create column that is growth
  bind_cols(Initial, Final) %>% 
    #binds initial (Diameter) and final (Diameter1) sizes to calculate growth
  mutate(Growth = ((Diameter1 - Diameter)/Diameter*100)) %>% 
    #add column for growth calculation
  select("Growth")


resultsGrow <-
  #table of initial and final size data
  results %>% 
    select("Day", "Treatment", "Temperature", "pH", "TankID", "Diam1", "Diam2") %>% 
    drop_na(Diam1, Diam2) %>% 
    gather(key = "1or2", value = "Diameter", "Diam1", "Diam2") %>% 
    filter(Day == "126")
resultsGrow <- bind_cols(resultsGrow, Growth)
#combines table of initial and final with growth information

#Figure out means of each treatment group
resultsGrow %>% 
  dplyr::group_by(Temperature, pH) %>% 
  dplyr::summarise(mean = mean(Growth), s.e. = se(Growth))

```

##### Growth linear mixed model results (Day 1 as initial)
```{r, GrowthResults=TRUE}
#LMM for growth:
modGrow <-
  resultsGrow %>% 
  lmer(Growth~ Temperature * pH + (1|TankID), data=.)

#Need to decide on model type that we want to use. 

summary(modGrow)

anova(modGrow, type="II") #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type II Analysis of Variance Table with Satterthwaite's method.
```

##### Check assumptions for growth model (Day 1 as initial)  
```{r, results=TRUE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modGrow)) 
shapiro.test(residuals(modGrow)) #passes
hist(residuals(modGrow)) #looks normal


# 2. Equal variances
leveneTest(residuals(modGrow)~resultsGrow$Treatment) #Passes
plot(fitted(modGrow),resid(modGrow,type="pearson"),col="blue")
```

##### Display exploratory interaction plot for growth model (Day 1 as initial)  
```{r, results=TRUE}
#Interaction plot shows slight interaction
interaction.plot(resultsGrow$Temperature,resultsGrow$pH,resultsGrow$Growth,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(Growth)),)
```


## Calcification


### Distal End (Tips of spines)    
##### Assemble data for chosen spine tips
```{r Calcification tip, results=FALSE}
### Model 1: calcification at the tips of spines

resultsTip <- 
  #create data using only the SEM images of the tips of spines.
  results %>% 
  select ("Treatment", "Temperature", "pH", "PartOfSpine", "Chosen", "RatioSEM", "TankID") %>% 
    #selects for desired variables. PartOfSpine is tip or base, Chosen indicates which side of the image (arbitrary) is selected for (aka has no dust)
  drop_na(RatioSEM) %>% 
  filter(Chosen == "yes", PartOfSpine == "Tip")
    #filters only for tips and chosen side. No dusty images to be analyzed.
```

##### Results for spine tips with linear mixed model
```{r, results=TRUE}
#LMM for calcification at the tips of spines 
modTip <-
  resultsTip %>% 
  lmer(RatioSEM ~ Temperature * pH + (1|TankID), data = .)

summary(modTip)

anova(modTip, type="II") #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type II Analysis of Variance Table with Satterthwaite's method.
```

#### Check assumptions for calcification at tip model.  
```{r, results=TRUE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modTip)) #Normal
shapiro.test(residuals(modTip)) #Pass
hist(residuals(modTip))

# 2. Equal variances
leveneTest(residuals(modTip)~resultsTip$Treatment) #Pass
plot(fitted(modTip),resid(modTip,type="pearson"),col="blue") 
```

#### Conduct post hoc analysis on calcification at tip.  
```{r, results=TRUE}
## POST HOC ANALYSIS 
emmTip <- emmeans(modTip, ~Temperature*pH, adjust = "tukey")  #Estimated marginal means (Least-squares means)
multcomp::cld(emmTip) #create a compact letter display of all pair-wise comparisons
pwpp(emmTip) #Constructs a plot of P values associated with pairwise comparisons of estimated marginal means.
```  

#### Display interaction plot for exploration of calcification at tip.  
```{r, results=TRUE}
#Interaction plot for cacification at the tip of the spines - interaction
interaction.plot(resultsTip$Temperature,resultsTip$pH,resultsTip$RatioSEM,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(Growth)),)
```


### Proximal End (Base of spines)  
#### Assemble data for chosen spine bases
```{r Calcification base}

### Model 2: calcification in the base of the spines

resultsBase <-
   #create data using only the SEM images of the base of spines.
  results %>% 
  select ("Treatment", "Temperature", "pH", "PartOfSpine", "Chosen", "RatioSEM", "TankID") %>% 
    #selects for desired variables. PartOfSpine is tip or base, Chosen indicates which side of the image (arbitrary) is selected for (aka has no dust)
  drop_na(RatioSEM) %>% 
  filter(Chosen == "yes", PartOfSpine == "Base")
    #filters only for tips and chosen side. No dusty images to be analyzed.
```

#### Results for spine bases with linear mixed model
```{r, results=TRUE}
#LMM for calcification at the tips of spines 
modBase <- 
  resultsBase %>% 
  lmer(RatioSEM~ Temperature * pH + (1|TankID), data=.)

summary(modBase) 
anova(modBase, type="II") #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.
```

#### Check assumptions for calcification at base model. 
```{r, results=TRUE}
## ASSSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modBase)) #Normal
shapiro.test(residuals(modBase)) # Pass
hist(residuals(modBase))

# 2. Equal variances
leveneTest(residuals(modBase)~resultsBase$Treatment) #Pass
plot(fitted(modBase),resid(modBase,type="pearson"),col="blue")
```

#### Conduct post hoc analysis on calcification at base  
```{r, results=TRUE}
## POST HOC ANALYSIS 
emmBase <- emmeans(modBase, ~Temperature*pH, adjust = "tukey")
  #Estimated marginal means (Least-squares means)
multcomp::cld(emmBase)
  #create a compact letter display of all pair-wise comparisons
pwpp(emmBase)
  #Constructs a plot of P values associated with pairwise comparisons of estimated marginal means.
```

#### Display interaction plot for exploration of calcification at base  
```{r, results=TRUE}
#Interaction plot for calcification at the base of the spines - no interaction
interaction.plot(resultsBase$Temperature,resultsBase$pH,resultsBase$RatioSEM,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(RatioSEM)),)

```


## Dropped Spines


##### Assemble data for dropped spines  
```{r Dropped Spines}

## Number of loose spines counted at the bottom of tanks at the end of the experiment. These were either broken or shed. Note: these spines were not collected, just counted, so can not determine whether they were broken or shed.

resultsDropped <-
  #create data using only the needed variables.
  results %>% 
  select ("Treatment", "Temperature", "pH", "SpineCount", "TankID") %>% 
  drop_na(SpineCount)
```

##### Analyse dropped spines with linear mixed effect model.  
```{r, results=TRUE}
##LM of dropped spines - can't make it LMM because error: number of levels of each grouping factor must be < number of observations - only have one count for each TankID.
modDropped <- 
  resultsDropped %>% 
  lm(SpineCount~ Temperature * pH, data=.)

summary(modDropped)
anova(modDropped) #anova{stats}.Compute analysis of variance (or deviance) tables for one or more fitted model objects. Produces Type III Analysis of Variance Table with Satterthwaite's method.
```

##### Check assumptions for model for dropped spines
```{r, results=TRUE}
##ASSUMPTIONS
# 1. Normality of residuals
qqPlot(residuals(modDropped)) #not normal
shapiro.test(residuals(modDropped)) #faaaiiilll
hist(residuals(modDropped))

# 2. Equal variances
leveneTest(residuals(modDropped)~resultsDropped$Treatment) # barely pass
plot(fitted(modDropped),resid(modDropped,type="pearson"),col="blue")
```

##### Analysis for dropped spines does not meet assumptions. Attempt a square root transformation.  
```{r, results=TRUE}
##DOES NOT MEET ASSUMPTIONS so ATTEMPT AT Transformation:
resultsDropped$tdata<-(resultsDropped$SpineCount)^1/2
  # transformation
modDropped1 <- lm(tdata~ Temperature * pH, data=resultsDropped)
  # run lm model of transformed data

###ASSUMPTIONS for Transformation
# 1. Normality of residuals
qqPlot(residuals(modDropped1)) #not normal
shapiro.test(residuals(modDropped1)) #fail
hist(residuals(modDropped1))
# 2. equal variances
bartlett.test(residuals(modDropped1)~resultsDropped$Treatment) #fail
plot(fitted(modDropped1),resid(modDropped1,type="pearson"),col="blue")


summary(modDropped1)
Anova(modDropped1, type=2) #test for significance of model
```

##### Transfomraiton (of any usual kind) of dropped spines was not successful, so we use a non parametric Kruskal Wallis Test.  
```{r, results=TRUE}
### Transformations not work, so use non parametric:
kruskal.test(resultsDropped$SpineCount~resultsDropped$Treatment)# SIG*** (p=0.0005563)
```

##### Conduct Dunn Post Hoc Test for dropped spines
```{r, results=TRUE}
## POST HOC Pairwise
dunn <- 
  resultsDropped %>% 
  dunnTest(SpineCount ~ Treatment,
           method = "holm", kw = TRUE, data = .)
dunnph <- dunn$res
cldList(P.adj ~ Comparison, data = dunnph, threshold = 0.05)
```

##### Display interaction plot for dropped spines 
```{r, results=TRUE}
#Interaction plot for dropped spines - no interaction
interaction.plot(resultsDropped$Temperature,resultsDropped$pH,resultsDropped$SpineCount,
                 type = c("l","p", "b", "o","c"), legend = TRUE,
                 trace.label = deparse(substitute(pH)),
                 fixed = FALSE,
                 xlab = deparse(substitute(Temperature)),
                 ylab = deparse(substitute(SpineCount)),)

```

