---
title: "Plots for growth and calcification responce of the common collector urchin, _Tripneustes gratilla_, to projected climate change: effects of warming and acidification."
author: "Emily Sesno"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
    smart: false
  pdf_document:
    keep_tex: yes
---

```{r setup, message=FALSE, warning=FALSE}
rm(list=ls(all=TRUE))
results <- read.csv("Data/Urchin_ClimateChange_Data.csv", header=T, na.strings="NA")

library("tidyverse")
library("lme4") #linear mixed modeling
library("lmtest") #linear mixed modeling
library("lmerTest") #calculate p-values in models
library("car") #levenes tests
library("emmeans") #post-hoc tests
library("lsmeans") #post-hoc tests
library("plotrix") #calculate std.error
library("dplyr") 
library("cowplot") #grid plotting 
```


### Treatment Conditions

```{r, results=TRUE}
#temp between high and ambient temp
summary(aov(Temp ~ Temperature, data=results)) #temperature was different between high and ambient treatment

#pH between high and ambient co2
summary(aov(pHspec ~ pH, data=results)) #pH was different between high and ambient co2 treatments

#temp between headers in high
aov1<- results %>%
  filter(Temperature=="heated")

summary(aov(Temp ~ as.factor(Header), data=aov1)) #not different between headers

#temp between headers in ambient
aov2<- results %>%
  filter(Temperature=="ambient")

summary(aov(Temp ~ as.factor(Header), data=aov2)) #not different between headers

#pH between headers in high
aov3<- results %>%
  filter(pH=="acidified")

summary(aov(pHspec ~ as.factor(Header), data=aov3)) #not different between headers

#pH between headers in ambient
aov4<- results %>%
  filter(pH=="ambient")

summary(aov(pHspec ~ as.factor(Header), data=aov4)) #not different between headers

```

```{r, results=TRUE}
tempsummary<-results %>% 
  select("Day", "Temperature", "Temp") %>%
  drop_na(Temp) %>%
  group_by(Day, Temperature) %>%
  mutate(meanT = mean(Temp)) %>%
  group_by(Temperature) %>% 
  mutate(sd = sd(Temp)) %>%
  mutate(se = sd/sqrt(12)) 

tempplot<-ggplot(data=tempsummary, aes(Day, meanT, color = Temperature)) +
    geom_point(size = 2.5, show.legend=FALSE) +
    geom_errorbar(aes(ymin = meanT-se, ymax = meanT+se)) +
    geom_line(size=1.2) +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="Temperature (°C)", breaks = seq(20,30, by = 1)) +
    scale_shape_discrete(name = NULL,
                         labels = c("Ambient", "High"))+
    scale_color_manual(name = NULL,
                       values = c("gray40", "firebrick3"),
                       labels = c("Ambient", "High")) +
    ggtitle("")+
    theme_classic() +
    theme(plot.margin=unit(c(0.3,0.6,0.3,0.3), "cm"))+
    theme(legend.margin = margin(0), 
          legend.position = c(.98,.8), 
          legend.justification = c("right", "top"),
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14, face="bold"));tempplot

#ggsave(filename="Figures/20200611/TempFig1.png", plot=tempplot, dpi=300, width=7, height=5, units="in")

```

```{r, results=TRUE}

pHsummary<-results %>% 
  select("Day","pH", "pHspec") %>%
  drop_na(pHspec) %>%
  group_by(Day, pH) %>%
  mutate(meanpH = mean(pHspec)) %>%
  group_by(pH, meanpH) %>% 
  mutate(sd = sd(pHspec)) %>%
  mutate(se = sd/sqrt(12)) 
  
  
pHplot<-ggplot(data=pHsummary, aes(Day, meanpH, color = pH)) +
    geom_point(size = 2.5, show.legend=FALSE) +
    geom_errorbar(aes(ymin = meanpH-sd, ymax = meanpH+sd)) +
    geom_line(size=1.2) +
    scale_x_continuous(name="Time (Day)", breaks = seq(0,130, by = 10)) +
    scale_y_continuous(name="pH (Total Scale)", breaks = seq(7.5,8.1, by = .1)) +
    scale_shape_discrete(name = NULL,
                         labels = c("Low pH", "High pH"))+
    scale_color_manual(name = NULL,
                       values = c("skyblue3", "gray40"),
                       labels = c("Acidified", "Ambient"),
                       guide = guide_legend(reverse = TRUE)) +
    ggtitle("")+
    theme_classic() +
    theme(plot.margin=unit(c(0.3,0.6,0.3,0.3), "cm"))+
    theme(legend.margin = margin(0), 
          legend.position = c(.98,.8), 
          legend.justification = c("right", "top"),
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14, face="bold"));pHplot

#ggsave(filename="Figures/20200611/pHFig1.png", plot=pHplot, dpi=300, width=7, height=5, units="in")

```


```{r, results=TRUE}

#Make plots in horizontal line with proper labeling of a and b

treatmentfig1<-plot_grid(tempplot, pHplot, labels = c("a", "b"), ncol=2, nrow=1, rel_heights= c(1,1), rel_widths = c(1,1), label_size = 20, label_y=1, align="h");treatmentfig1

#ggsave(filename="Figures/20200611/TreatmentFig1.png", plot=treatmentfig1, dpi=300, width=16, height=6, units="in")

```

**Figure 3.1.** Temperature (a) and pH (b) conditions over the 126-day experiment period. Data are mean ± standard error (s.e.). Where error bars are not visible, error is too small to be seen. Days prior to 0 indicate the acclimation and conditioning period.


### Growth: 


```{r, echo = FALSE}
#### Growth Interaction Plot
##May need to adjust the p value for temp depending on which Anova() or anova() we use...

growthsummary<-results %>% 
  filter(Day==126) %>%
  select("pH", "Temperature","Growth") %>%
  drop_na(Growth) %>% 
  mutate(meanPct = Growth*100) %>%
  group_by(pH, Temperature) %>%
  summarise(mean = mean(meanPct), 
            N = length(meanPct),
            se = std.error(meanPct))
   
  
growthplot<-ggplot(data=growthsummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Total Growth (%)")+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    geom_text(x=1.5, y=125, label="p(pH)=0.61", size=6, color="darkgray") + 
    geom_text(x=1.5, y=100, label="p(Temperature)=0.06", size=6, color="darkgray") + 
    geom_text(x=1.5, y=75, label="p(Interaction)=0.47", size=6, color="darkgray") + 
    ylim(0,450)+
    theme_classic()+
    theme(legend.position = "right", 
          legend.background = element_blank(),
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));growthplot

#ggsave(filename="Figures/20200611/GrowthFig.png", plot=growthplot, dpi=300, width=8, height=6, units="in")

```

**Figure 3.2.** Effect of (a) temperature and (b) pH on growth (%) of Tripneustes gratilla test diameters after 126-day exposure period. Data are mean ± standard error (s.e.), n= 11-12.


### Calcification Ratio  

```{r, echo = FALSE}
#### Tip and base calcification ratios

#base
basesummary<-results %>% 
  select("Temperature","pH","RatioSEM", "PartOfSpine") %>%
  filter(PartOfSpine=="Base")%>%
  drop_na(RatioSEM) %>% 
  group_by(pH, Temperature, PartOfSpine) %>%
  summarize(mean = mean(RatioSEM), se = std.error(RatioSEM), N = length(RatioSEM))
  
baseplot<- ggplot(data=basesummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Spine Base Calcification Ratio")+
    ylim(1,3.5)+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    geom_text(x=1.5, y=1.6, label="p(pH)=0.002", size=6, color="black") + 
    geom_text(x=1.5, y=1.4, label="p(Temperature)=0.16", size=6, color="darkgray") + 
    geom_text(x=1.5, y=1.2, label="p(Interaction)=0.54", size=6, color="darkgray") + 
    theme_classic()+
    theme(legend.position = "right", 
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));baseplot

#ggsave("CalcRatiobase.png", plot=baseplot, path = "/Users/emilysesno/Desktop/R_Analysis/R_Analysis/output",  width = 7, height = 5)  

#ggsave(filename="Figures/20200611/BaseCalcificationFig.png", plot=baseplot, dpi=300, width=8, height=6, units="in")

#tip
tipsummary<-results %>% 
  select("Temperature","pH","RatioSEM", "PartOfSpine") %>%
  filter(PartOfSpine=="Tip")%>%
  drop_na(RatioSEM) %>% 
  group_by(pH, Temperature, PartOfSpine) %>%
  summarize(mean = mean(RatioSEM), se = std.error(RatioSEM), N = length(RatioSEM))

tipplot<-ggplot(data=tipsummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Spine Tip Calcification Ratio")+
    ylim(1,3.5)+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    geom_text(x=1.5, y=3.0, label="p(pH)=0.44", size=6, color="darkgray") + 
    geom_text(x=1.5, y=2.8, label="p(Temperature)=0.39", size=6, color="darkgray") + 
    geom_text(x=1.5, y=2.6, label="p(Interaction)=0.09", size=6, color="darkgray") + 
    theme_classic()+
    theme(legend.position = "none", #removed the legend for the tip, so when we align them horizontally we only have one legend. But need to adjust the size so it is the same as base
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));tipplot


#ggsave(filename="Figures/20200611/TipCalcificationFig.png", plot=tipplot, dpi=300, width=8, height=6, units="in")
```

```{r, results=TRUE}

#Make plots of calcification at tip and base in horizontal line with proper labeling of a and b
#removed legend on the tip, so they can share the same one. But now they are different sizes...

calcificationfig1<-plot_grid(tipplot, baseplot, labels = c("a", "b"), ncol=2, nrow=1, rel_heights= c(1,1), rel_widths = c(0.8,1), label_size = 20, label_y=1, align="h");calcificationfig1

#ggsave(filename="Figures/20200611/CalcificationFig1.png", plot=calcificationfig1, dpi=300, width=16, height=6, units="in")

```


**Figure 3.3.** Effect of temperature and pH on the calcification ratio at the tip (a) and base (b) of spine cross-sections. Data are means ± standard error (s.e.), n = 14-18.




### Spine Loss 

```{r, echo = FALSE}

droppedsummary<-results %>% 
  select("Temperature","pH","SpineCount") %>%
  drop_na(SpineCount) %>% 
  group_by(pH, Temperature) %>%
  summarize(mean = mean(SpineCount), se = std.error(SpineCount), N = length(SpineCount))

  dropplot<-ggplot(data=droppedsummary, aes(x=Temperature, y=mean, color = pH, group=interaction(Temperature,pH))) +
    geom_point(size = 3,  position=position_dodge(0.1)) +
    geom_line(aes(group=pH),  position=position_dodge(0.1)) +
    xlab("Temperature Treatment") +
    ylab("Dropped Spines")+
    geom_errorbar(aes(ymin = mean-se, ymax = mean+se, width = 0.1), position=position_dodge(0.1)) +
    ylim(0,30) +
    theme_bw() + 
    scale_x_discrete(limits=c("ambient", "heated"), 
                     label=c("Ambient", "High"))+
    scale_colour_manual(name = "pH Treatment",
                        values = c("skyblue3", "gray40"),
                        labels = c("Acidified", "Ambient"),
                        guide = guide_legend(reverse = TRUE)) +
    #geom_text(x=1.5, y=11, label="p(pH)<0.001", size=6, color="black") + 
    #geom_text(x=1.5, y=9, label="p(Temperature)=0.91", size=6, color="darkgray") + 
    #geom_text(x=1.5, y=7, label="p(Interaction)=0.83", size=6, color="darkgray") +
    theme_classic()+
    theme(legend.position = "right", 
          legend.background = element_blank(), 
          axis.title = element_text(size = 16, face="bold"),
          axis.text = element_text(size = 14), 
          plot.title = element_text(size=16, face="bold"),
          legend.text = element_text(size=14), 
          legend.title = element_text(size=16, face="bold"));dropplot

#ggsave(filename="Figures/20200611/DroppedSpinesFig.png", plot=dropplot, dpi=300, width=8, height=6, units="in")
```

**Figure 3.5.** Effect of temperature and pH on the number of spines dropped. Data are means + standard error (s.e.), n = 5-6  
